\chapter{Corollary Projects}
\label{cha:corollary_projects}

During the development, I noticed that several portions of the code might be
turned into independent libraries and utility scripts that may be valuable to
other programmers as well as my single use-case scenario. Even though it is now external
to reCluster, this software is still an essential part of it, which is now open to
the entire community. \\ %
Shortly after the publication, I began receiving some feedback, contributions,
and, most importantly, appreciation in the form of GitHub stars\footnote{\url{https://docs.github.com/en/get-started/exploring-projects-on-github/saving-repositories-with-stars}}.
\\ %
% TODO Reference Philosophy and MIT
The three projects derived from the development of reCluster are briefly
illustrated and explained in the sections that follow. As stated in the Philosophy
section, everything is completely Open Source and available under the MIT
license.

\section{Node Exporter Installer}
\label{sec:corollary_projects_node_exporter_installer}

% TODO Node exporter reference
% TODO K3s reference
Available at \url{https://github.com/carlocorradini/node\_exporter\_installer} \\ %
Used while installing reCluster on a Node. \\ %
Node exporter does not provide any installation script and the default procedure
(see \url{https://github.com/prometheus/node_exporter#installation-and-usage})
is far from user-friendly and easily configurable. \\ %
Inspired by K3s \texttt{install.sh}\footnote{\url{https://github.com/k3s-io/k3s/blob/master/install.sh}}
script, Node exporter installer helps the user by automatically installing Node
exporter on the machine. Condensed in a single \texttt{install.sh} POSIX script,
it is easily configurable (see section
\ref{subsec:corollary_projects_node_exporter_installer_configuration}) and can be
downloaded and executed directly by \texttt{sh} (see example \ref{subsubsec:corollary_projects_node_exporter_installer_example_basic}).

\subsection{Configuration}
\label{subsec:corollary_projects_node_exporter_installer_configuration}

Node exporter installer accepts environment variables only as configuration parameters.
This is done to avoid any potential conflict with the default behavior of Node
exporter, which employs argument flags. See
\url{https://github.com/prometheus/node\_exporter\#collectors} for a
comprehensive list of Node exporter configuration parameters. \\ %
The configuration settings allowed by Node exporter installer are shown in the
table below.

\begin{xltabular}
  {\textwidth} { >{\ttfamily}l | X | >{\ttfamily}c }

  \multicolumn{1}{ c |}{\large{\textbf{Name}}} &
  \multicolumn{1}{ c |}{\large{\textbf{Description}}} &
  \multicolumn{1}{ c }{\large{\textbf{Default Value}}} \\ \hhline{===}

  INSTALL\_NODE\_EXPORTER\_SKIP\_DOWNLOAD & Skip downloading Node exporter.
  \newline
  A local executable binary must already exist at \texttt{<BIN\_DIR>/node\_exporter}
  \newline
  Useful in an Air-Gapped environment. % TODO Air-Gap environment reference
  & false \\ \hline

  INSTALL\_NODE\_EXPORTER\_FORCE\_RESTART & Force restarting Node exporter service.
  & false \\ \hline

  INSTALL\_NODE\_EXPORTER\_SKIP\_ENABLE & Do not enable Node exporter service at
  startup. & false \\ \hline

  INSTALL\_NODE\_EXPORTER\_SKIP\_START & Do not start Node exporter service. &
  false \\ \hline

  INSTALL\_NODE\_EXPORTER\_SKIP\_FIREWALL & Do not apply any firewall rules.
  \newline
  Supported firewalls are: \texttt{iptables}\footnote{\url{https://www.netfilter.org/projects/iptables}},
  \texttt{firewalld}\footnote{\url{https://firewalld.org}} and \texttt{UFW}\footnote{\url{https://wiki.ubuntu.com/UncomplicatedFirewall}}.
  & false \\ \hline

  INSTALL\_NODE\_EXPORTER\_SKIP\_SELINUX & Do not change \texttt{SELinux\footnote{\url{https://selinuxproject.org}}}
  context for Node exporter binary. & false \\ \hline

  INSTALL\_NODE\_EXPORTER\_VERSION & Node exporter version to download. & latest
  \\ \hline

  INSTALL\_NODE\_EXPORTER\_BIN\_DIR & Directory where to install Node exporter
  binary and uninstall script. & \makecell[Xt]{\centering{/usr/local/bin \\ or \\ /opt/bin}}
  \\ \hline

  INSTALL\_NODE\_EXPORTER\_SYSTEMD\_DIR & Directory where to install Systemd
  service files. & /etc/systemd/system \\ \hline

  INSTALL\_NODE\_EXPORTER\_EXEC & Node exporter configuration flags. & \\

  \caption{Node exporter installer configuration parameters}
\end{xltabular}

\subsection{Example}
\label{subsec:corollary_projects_node_exporter_installer_example}

This section shows some examples of how to use Node exporter installer. In both,
we use \texttt{curl} (a command-line tool for getting or sending data using URL
syntax\cite{curl}) to download the script from
\url{https://raw.githubusercontent.com/carlocorradini/node_exporter_installer/main/install.sh}
and pipe (redirect the output of a program to the input of another program) the
result to \texttt{sh} (UNIX shell command-line interpreter\footnote{\url{https://wikipedia.org/wiki/Unix_shell}})
for execution.

\subsubsection{Basic}
\label{subsubsec:corollary_projects_node_exporter_installer_example_basic}

Install Node exporter using the default configuration parameters for both
installer and binary.

\begin{lstlisting}[language=shell, morekeywords={[3]{--silent, --show-error, --fail, --location, -}}, xleftmargin=\parindent, caption=Basic installation with default configuration parameters]
  curl \
    --silent \      # Do not show progress meter or error messages
    --show-error \  # Show an error message if it fails
    --fail \        # Fail fast with no output at all on server errors
    --location \    # If the requested resource has moved, redo the request to the new location
    "https://raw.githubusercontent.com/carlocorradini/node_exporter_installer/main/install.sh" \
    | sh -
\end{lstlisting}

\subsubsection{Advanced}
\label{subsubsec:corollary_projects_node_exporter_installer_example_advanced}

Install Node exporter v1.5.0 without starting the service automatically. Disable
all default collectors, leaving only CPU and memory statistics enabled.

\begin{lstlisting}[language=shell, alsoletter={-, .}, morekeywords={[3]{..., --collector.disable-defaults, --collector.cpu, --collector.meminfo, -}}, morekeywords={[4]{INSTALL_NODE_EXPORTER_VERSION, INSTALL_NODE_EXPORTER_SKIP_START}}, xleftmargin=\parindent, caption=Advanced installation with custom configuration parameters]
  curl \
    ... \
    | INSTALL_NODE_EXPORTER_VERSION="v1.5.0" \   # Download and install Node exporter v1.5.0
      INSTALL_NODE_EXPORTER_SKIP_START="true" \  # Do not start Node exporter service
      sh - \
        --collector.disable-defaults \             # Disable default collectors
        --collector.cpu \                          # Expose CPU statistics
        --collector.meminfo                        # Expose memory statistics
\end{lstlisting}

\section{Inline}
\label{sec:corollary_projects_inline}

% TODO Bundle reference
Available at \url{https://github.com/carlocorradini/inline} \\ %
Used while generating the final release bundle for distribution. \\ %
It is useful to be able to split a large script into many files to make it easier
to work with while still being able to distribute it as a single script. This
program reads an input file and produces an output file with all of the sources
inlined. \\ %
The source command, \texttt{.\ <FILE>} for POSIX shells or \texttt{source <FILE>}
for non-POSIX shells (e.g. Bash\footnote{\url{https://www.gnu.org/software/bash}}),
read, and executes commands from the file specified in the current shell
environment. It is useful to load functions, variables, and configuration files into
the current shell context\cite{source}. \\ %
Inline is a static tool that does not execute the input script. As a result, it
cannot determine the value of a variable dynamically if it is used inside the
source command path (i.e. \texttt{source "\$DIR/path/to/script/sh"}). To avoid
this, inline requires a hint on where to find the specified file.

\subsection{Features}
\label{subsec:corollary_projects_inline_features}

Many helpful features of Inline are described below.

\begin{itemize}
  \item POSIX standard-compliant.

  \item Sourcing with quotes, spaces, and more.

  \item Sourcing from global variable \texttt{\$PATH}.

  \item Sourcing from \texttt{ShellCheck} (a shell script static analysis tool\footnote{\url{https://www.shellcheck.net}})
    source directive\footnote{\url{https://www.shellcheck.net/wiki/Directive}}.
    \newline
    This is considered a hint to Inline and is used only if the source path is invalid.
    \newline
    \begin{lstlisting}[language=shell, alsoletter={.}, morekeywords={[2]{.}}, numbers=none, aboveskip=0pt, belowskip=0pt, abovecaptionskip=0pt, belowcaptionskip=0pt]
      # shellcheck source=path/to/script.sh
      . "$DIR/path/to/script.sh"

      # shellcheck source=path/to/script.sh
      source "$DIR/path/to/script.sh"
    \end{lstlisting}

  \item Recursive sources. If a sourced file contains additional source commands,
    they are also inlined and included in the final output script.
    \newline
    It should be noted that these can cause infinite recursion.

  \item Recursion detection. To avoid infinite recursion, an exception is thrown
    if a file is sourced multiple times. This is accomplished through the use of
    an internal cache that saves the absolute path to each sourced file. If the path
    to a script file in a source command is already in the cache, a recursion is
    detected.

  \item Shebang removal in sourced files.
    \newline
    A shebang is the character sequence consisting of the characters number sign
    and exclamation mark (\texttt{\#!}) at the beginning of a script. In a Unix-like
    operating system, when a text file with a shebang is used as if it is an
    executable, the program loader executes the specified interpreter program,
    passing as an argument the path that was initially used when attempting to
    run the script, so that the program can use the file as input data\cite{shebang}
    \newline
    The shell interpreter only allows one shebang per script. As a result, the
    shebang is only allowed in the input script file, while it is automatically removed
    in all sourced script files.

  \item Avoid inlining a certain source file. If the inline skip comment
    directive (\texttt{\# inline skip}) is present before a source command, the
    latter is ignored and not inlined. As a consequence, the final output script
    includes the original command, unaltered and unaligned.
    \newline
    It is worth noting that the skip directive also works with ShellCheck. The sole
    requirement is that there are no blank lines between the directives and the source
    command.
    \newline
    \begin{lstlisting}[language=shell, alsoletter={.}, morekeywords={[2]{.}}, numbers=none, aboveskip=0pt, belowskip=0pt, abovecaptionskip=0pt, belowcaptionskip=0pt]
    # inline skip
    # shellcheck source=path/to/script.sh
    . "$DIR/path/to/script.sh"

    # shellcheck source=path/to/script.sh
    # inline skip
    source "$DIR/path/to/script.sh"
  \end{lstlisting}
\end{itemize}

\subsection{Configuration}
\label{subsec:corollary_projects_inline_configuration}

Inline's behavior is easily customizable by using argument flags, which begin with
a double dash followed by the property name and an optional value (\texttt{--NAME
[VALUE]}). \\ %
The accepted configuration parameters are listed in the table below.

\begin{xltabular}
  {\textwidth} { >{\ttfamily}l | X | >{\ttfamily}c }

  \multicolumn{1}{ c |}{\large{\textbf{Name}}} &
  \multicolumn{1}{ c |}{\large{\textbf{Description}}} &
  \multicolumn{1}{ c }{\large{\textbf{Default Value}}} \\ \hhline{===}

  --in-file <FILE> & Input script file (\texttt{<FILE>}).
  \newline
  If the file does not exist, an error is thrown. & \\ \hline

  --out-file <FILE> & Output script file (\texttt{<FILE>}).
  \newline
  If a file with the same name and path already exists, an error is thrown. Unless
  the \texttt{--overwrite} argument is given, the latter is always true (see
  option below).
  \newline
  If no value is specified, the name of the output file is determined by examining
  the original input file. The first portion is the original name, followed by a
  \texttt{.inlined} string, and finally, if present, the extension (usually \texttt{.sh}).
  & <NAME>.inlined[EXTENSION] \\ \hline

  --overwrite & Replace the input file.
  \newline
  The inlined result replaces the original file content. This is the same as setting
  the output file's value as the original input file, except that the existence
  check is bypassed (see option above). & false \\ \hline

  --log-level <LEVEL> & Logging level (\texttt{<LEVEL>}).
  \newline
  Attachment \ref{cha:logging} provides additional information regarding logging
  and logging levels.
  \newline
  The following logging levels are supported (listed in descending order of
  importance):
  \begin{itemize}[noitemsep]
    \item[\protect\icircled{\texttt{5}}] \texttt{silent}

    \item[\protect\icircled{\texttt{4}}] \texttt{fatal}

    \item[\protect\icircled{\texttt{3}}] \texttt{warn}

    \item[\protect\icircled{\texttt{2}}] \texttt{info}

    \item[\protect\icircled{\texttt{1}}] \texttt{debug}
  \end{itemize}
  & info \\ \hline

  --disable-color & Disable terminal colors (enabled by default). & false \\
  \hline

  --help & Display a help message and terminate (successfully). & \\

  \caption{Inline configuration parameters}
\end{xltabular}

\subsection{Example}
\label{subsec:corollary_projects_inline_example}

This section provides an example of how Inline works and how it may be used. \\ %
There are two script files, both of which begin with a shebang. There is a print
command in the first file, \texttt{hello.sh}, that outputs \texttt{Hello} and an
empty space, followed by a source command that includes the script file \texttt{world.sh}.
The format of the second file, \texttt{world.sh}, is the same as the first, except
that the print command outputs \texttt{World!} and the newline character (\texttt{\char`\\n}),
and there are no source commands. \\ %
The following two listings show the contents of the two files:

\noindent
\hspace{.775\parindent}
\begin{minipage}[t]{.45\textwidth}
  \begin{lstlisting}[language=shell, alsoletter={.}, morekeywords={[2]{.}}, caption=Input script \texttt{hello.sh}]
    #!/usr/bin/env sh

    printf "Hello "

    . "world.sh"
  \end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}[t]{.45\textwidth}
  \begin{lstlisting}[language=shell, alsoletter={.}, morekeywords={[2]{.}}, caption=Sourced script \texttt{world.sh}]
    #!/usr/bin/env sh

    printf "World!\n"
  \end{lstlisting}
\end{minipage}

The goal is to inline the script file \texttt{hello.sh} and produce an output
file that has no source commands. When the command \texttt{./inline.sh --in-file
"hello.sh"} is executed, the following result is obtained:

\begin{lstlisting}[language=shell, alsoletter={.}, morekeywords={[2]{.}}, xleftmargin=\parindent, caption=Inlined script \texttt{hello.inlined.sh}]
  #!/usr/bin/env sh

  printf "Hello "

  # . "world.sh"

  printf "World!\n"
\end{lstlisting}

Note the unique shebang as well as the source command that has been commented
out with the symbol \texttt{\#}. Furthermore, because no \texttt{--out-file}
option is used, the final output file is named \texttt{hello.inlined.sh}, which does
not override the original input file.

\section{GraphQL Auth Directive}
\label{sec:corollary_projects_graphql_auth_diretive}

% TODO GraphQL reference
% TODO Node.js reference
Available at \url{https://github.com/carlocorradini/graphql-auth-directive} \\ %
Used in the \texttt{GraphQL} API. \\ %
A custom \texttt{GraphQL} directive that protects resources from unauthenticated
and unauthorized access in high-security contexts. It is available in all major
\texttt{Node.js} registries as \texttt{graphql-auth-directive}. \\ %
A directive is an identifier preceded by a \texttt{@} character, optionally
followed by a list of named arguments, which can appear after almost any form of
syntax in the \texttt{GraphQL} query or schema languages\cite{graphql_directive}.
\\ %
The \texttt{GraphQL} context holds all important information about the current
request. A \texttt{GraphQL} context is an object that is shared by all resolvers
in a given execution. It helps store data such as authentication information,
the current user, database connections, data sources, and other information required
to operate the business logic\cite{graphql_context}. It is important to note that
the context does not have to follow a predefined structure; rather, it is highly
flexible to the user's implementation.

\subsection{Configuration}
\label{subsec:corollary_projects_graphql_auth_diretive_configuration}

Before applying the directive to the \texttt{GraphQL} schema, it must be built/configured.
The library exposes a utility function called \texttt{buildAuthDirective} that
accepts a configuration object with the values provided in the table below and
returns the type definition (see section \ref{subsec:corollary_projects_graphql_auth_diretive_type_definition})
as well as a transformer function that applies the auth logic to the executable
\texttt{GraphQL} schema.

\begin{xltabular}
  {\textwidth} { >{\ttfamily}l | X | >{\ttfamily}c }

  \multicolumn{1}{ c |}{\large{\textbf{Name}}} &
  \multicolumn{1}{ c |}{\large{\textbf{Description}}} &
  \multicolumn{1}{ c }{\large{\textbf{Default Value}}} \\ \hhline{===}

  name & Directive name.
  \newline
  If a name different from the default is specified, it must be reflected in the
  schema where the directive is used, otherwise, an error is thrown. & auth \\ \hline

  auth & A function or class that handles authentication and authorization.
  \newline
  The current context (which contains information about the current request) and
  the roles and permissions required by the requested resource must be accepted as
  arguments by the implementation. If access to the requested resource is
  granted, the boolean value \texttt{true} is returned; otherwise, \texttt{false}
  is returned if access is denied.
  \newline
  A default basic auth function is already implemented and checks for the
  existence of an authorized applicant and that its roles and permissions
  overlap with those of the requested resource. & \\ \hline

  authMode & Auth mode if access is not granted.
  \newline
  Methodology for informing the requestor that access to the resource has been denied.
  It is often desirable to hide the important information that the request is
  failing due to an auth check and instead deliver an error/informational
  response stating that the requested resource simply does not exist.
  \newline
  The first mode, \texttt{ERROR}, throws an authentication or authorization
  error (see below), whereas the second mode, \texttt{NULL}, returns the value
  \texttt{null} (empty). & ERROR \\ \hline

  roles & Roles configuration.
  \newline
  An object with two properties:
  \begin{itemize}
    \item \texttt{enumName}
      \newline
      Defines the array type, which is by default a \texttt{String}.
      \newline
      It is standard practice in \texttt{GraphQL} and in programming languages
      to map a set of values to an enum. An enumeration type is a special kind of
      scalar that is restricted to a particular set of allowed values\cite{graphql_enum}.
      \newline
      This option restricts the allowed values of roles to a specific set rather
      than a general string.

    \item \texttt{default}
      \newline
      Roles that are required by default.
      \newline
      No roles are required by default. As a result, access to a protected resource
      can only be provided with authentication. Overriding this option enables
      authorization by default, requiring the requestor to have at least one matching
      role.
  \end{itemize}
  & \makecell[Xt]{\centering{enumName: String \\ default: []}} \\ \hline

  permissions & Permissions configuration.
  \newline
  An object with the same properties as roles configuration (see above). &
  \makecell[Xt]{\centering{enumName: String \\ default: []}} \\ \hline

  authenticationError & Authentication error class. The error is thrown if there
  is no authenticated requestor in the current context. By default, a generic
  error is thrown. The class must extend the \texttt{Error} class. & AuthenticationError
  \\ \hline

  authorizationError & Authorization error class. The error is thrown if the roles
  and/or permissions of the requestor do not overlap with those of the requested
  resource. By default, a generic error is thrown. The class must extend the
  \texttt{Error} class. & AuthorizationError \\ \hline

  container & Dependency Injection Container.
  \newline
  Dependency injection is a design pattern that shifts the responsibility of
  resolving dependencies to a dedicated dependency injector that knows which
  dependent objects to inject into application code\cite{dependency_injection}.
  \newline
  It should be noted that dependency injection is only available if \texttt{auth}
  is a class type. & IOCContainer \\

  \caption{GraphQL auth directive configuration parameters}
\end{xltabular}

\subsection{Type Definition}
\label{subsec:corollary_projects_graphql_auth_diretive_type_definition}

% TODO GraphQL reference
GraphQL has its language for writing GraphQL Schemas: the GraphQL Schema
Definition Language (SDL). SDL is incredibly powerful and expressive while being
simple and intuitive to use. The syntax is well-defined and is included in the
official GraphQL specification\footnote{\url{https://spec.graphql.org}}. Schema Definitions
are also known as IDL (Interface Definition Language) or SDL (Schema Definition
Language)\cite{graphql_sdl}. \\ %
The Type Definition of the GraphQL auth directive generated with default
settings is shown below:

\begin{lstlisting}[language=graphql, morekeywords={[2]{@auth}}, morekeywords={[4]{roles, permissions}}, xleftmargin=\parindent, caption=GraphQL auth directive Type Definition]
  """
  Protect the resource from unauthenticated and unauthorized access.
  """
  directive @auth(
    """
    Allowed roles to access the resource.
    """
    roles: [String!]! = []
    """
    Allowed permissions to access the resource.
    """
    permissions: [String!]! = []
  ) on FIELD | FIELD_DEFINITION | OBJECT
\end{lstlisting}

The above definition must be available in every \texttt{GraphQL} schema that
utilizes the directive, otherwise an error is thrown during the server bootstrap
operation. \\ %
As stated in section \ref{subsec:corollary_projects_graphql_auth_diretive_configuration},
the definition is generated dynamically by a set of configuration options or
their default value. The name (\texttt{@auth}), \texttt{roles} and \texttt{permissions}
type (non-nullable \texttt{String} array), and default value (an empty array
\texttt{[]}) are easily distinguishable.

\subsection{Example}
\label{subsec:corollary_projects_graphql_auth_diretive_example}

Consider a \texttt{GraphQL} backend application that requires authentication and
authorization. The sections of the \texttt{GraphQL} schema where the auth directive
is used, as well as the enums that compose the different roles and permissions
mappings are listed below:

\begin{lstlisting}[language=graphql, morekeywords={[2]{@auth, Role, Permission}}, morekeywords={[3]{ADMIN, SIMPLE, VIEW, EDIT}}, morekeywords={[4]{roles, permissions, unprotected, protected, secret, arg}}, xleftmargin=\parindent, caption=\texttt{GraphQL} schema with auth directive and mappings]
  enum Role {
    ADMIN
    SIMPLE
  }

  enum Permission {
    VIEW
    EDIT
  }

  directive @auth(
    roles: [Role!]! = []
    permissions: [Permission!]! = []
  ) on FIELD | FIELD_DEFINITION | OBJECT

  type Query {
    unprotected: String!
    protected(arg: Boolean): Int! @auth
    secret: Float! @auth(roles: [ADMIN], permissions: [VIEW])
  }
\end{lstlisting}

The available roles are represented by the enum \texttt{Role}: \texttt{ADMIN}
for administrators and \texttt{BASIC} for simple users. The supported
permissions are represented by the enum \texttt{Permission}: \texttt{VIEW} to allow
resource visualization and \texttt{EDIT} to allow resource modification. \\ %
The definition of \texttt{@auth} directive accepts as roles an array of \texttt{Role}
whose default value is empty and as permissions an array of \texttt{Permission} whose
default value is empty. Notice the difference from the default type definition
shown in section
\ref{subsec:corollary_projects_graphql_auth_diretive_type_definition}. \\ %
Three queries are supported:
\begin{enumerate}
  \item \lstinline[language=graphql, basicstyle=\ttfamily, morekeywords={[4]{unprotected}}]{unprotected: String!}
    \newline
    \texttt{unprotected} does not accept any arguments and returns a non-nullable
    string. Does not require any authentication or authorization. As a result, it
    may be called by anyone.

  \item \lstinline[language=graphql, basicstyle=\ttfamily, morekeywords={[2]{@auth}},
    morekeywords={[4]{protected, arg}}]{protected(arg: Boolean): Int! @auth}
    \newline
    \texttt{unprotected} accepts a boolean optional parameter named \texttt{arg}
    and returns a non-nullable integer number. Because no roles or permissions are
    specified in the directive, it requires only authentication but no
    authorization. As a consequence, everyone who has been authenticated is
    allowed to call this query.

  \item \lstinline[language=graphql, basicstyle=\ttfamily, morekeywords={[2]{@auth}},
    morekeywords={[3]{ADMIN, VIEW}}, morekeywords={[4]{roles, permissions, secret}}]{secret: Float! @auth(roles: [ADMIN], permissions: [VIEW])}
    \newline
    The most secure of the three, \texttt{secret}, does not take any parameters
    and returns a non-nullable float integer. The caller must have an admin role
    with view permission to use this query. As a consequence, the query is secured
    by both authentication and authorization, resulting in higher requirements
    to perform the desired action on the resource.
\end{enumerate}