\chapter{Logging}
\label{cha:logging}

Logging is the process of preserving a record of events that occur in a computer
system, such as issues, errors, or just information on current operations. These
events might happen in the operating system or other applications. For each such
event, a message or log entry is recorded. These log messages may subsequently
be used to monitor and understand the system's operation, troubleshoot problems,
or during an audit. Logging is very crucial in multi-user applications to provide
a centralized view of the system's functionality\cite{logging}. \\ %
It is critical in the architecture implementation to log the numerous operations
that occur in each component of the cluster. The latter not only assists in understanding
how the general architecture performs and its status, but also the potential
causes of an incident. During the cluster's development, if any components
failed or did not behave as they should (e.g., no autonomous upscaling or downscaling),
monitoring and analyzing the numerous log files created was the only method to solve
the problems and understanding what and why happened. Furthermore, a logging system
is available for the multitude of utility scripts and applications that are not directly
employed with the cluster but rather during its general development. \\ %
Log interpretation is a challenging process. Log management system collects data
from a variety of sources with varying forms, purposes, and granularities. Maintaining
defined log levels helps give each entry meaning and better understand the significance
of the related log message\cite{logging_levels}. Section \ref{sec:logging_levels}
describes common logging levels that are also employed in cluster development and
implementation.

\section{Levels}
\label{sec:logging_levels}

A log level is a piece of information that indicates the importance of a log
message. It is a basic yet effective method of identifying log events from one another.
Furthermore, it is utilized to filter crucial information regarding the system state
to only those that are solely informational\cite{logging_levels}. \\ %
During the initialization phase, a level is chosen from the list below (sorted
in descending order of importance), with the expectation that the lower levels
are ignored in favor of the higher ones.

\begin{enumerate}
  \item[\protect\icircled{\texttt{7}}] \texttt{SILENT}
    \newline
    The highest possible logging level is meant to completely disable logging.

  \item[\protect\icircled{\texttt{6}}] \texttt{FATAL}
    \newline
    The system experienced a very severe error, which caused the system to abort.

  \item[\protect\icircled{\texttt{5}}] \texttt{ERROR}
    \newline
    Employed when the system encounters a problem that prevents one or more operations
    from functioning properly but allows it to continue operating.

  \item[\protect\icircled{\texttt{4}}] \texttt{WARN}
    \newline
    Designates potentially dangerous conditions that do not cause the system to crash.

  \item[\protect\icircled{\texttt{3}}] \texttt{INFO}
    \newline
    The standard log level indicates informational messages about the progress at
    a coarse-grained level. The information reported using the \texttt{INFO} log
    level should be strictly informative, without the need for any essential
    information to be lost.

  \item[\protect\icircled{\texttt{2}}] \texttt{DEBUG}
    \newline
    Used for information that may be required for diagnosing and debugging issues,
    or for running in a test environment to ensure that everything is working
    properly.

  \item[\protect\icircled{\texttt{1}}] \texttt{TRACE}
    \newline
    The most fine-grained level of information is only employed in rare circumstances
    where full visibility of what is occurring is required. \texttt{TRACE}'s logging
    level is quite verbose.
\end{enumerate}

Table \ref{tbl:logging_levels}, shows the correlation between the logging level
and the related logging message. The default logging level in various implementations
is often set to \texttt{INFO} level.

\begin{xltabular}
  {\textwidth} { >{\ttfamily}c | >{\large\ttfamily}c | >{\large\ttfamily}c | >{\large\ttfamily}c | >{\large\ttfamily}c | >{\large\ttfamily}c | >{\large\ttfamily}c }

  \multicolumn{1}{
  c
  |}{\backslashbox{\large{\textbf{Logging Level}}}{\large{\textbf{Logging Message}}}}
  & FATAL & ERROR & WARN & INFO & DEBUG & TRACE \\ \hhline{=======}

  \protect\icircled{\texttt{7}} SILENT & \cellcolor{bulmaRed}\hphantom{......} &
  \cellcolor{bulmaRed}\hphantom{......} & \cellcolor{bulmaRed}\hphantom{......}
  & \cellcolor{bulmaRed}\hphantom{......} & \cellcolor{bulmaRed}\hphantom{......}
  &\cellcolor{bulmaRed}\hphantom{......} \\ \hline

  \protect\icircled{\texttt{6}} FATAL\leavevmode\hphantom{.} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaRed} & \cellcolor{bulmaRed} & \cellcolor{bulmaRed} &
  \cellcolor{bulmaRed} & \cellcolor{bulmaRed} \\ \hline

  \protect\icircled{\texttt{5}} ERROR\leavevmode\hphantom{.} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaGreen} & \cellcolor{bulmaRed} & \cellcolor{bulmaRed} & \cellcolor{bulmaRed}
  & \cellcolor{bulmaRed} \\ \hline

  \protect\icircled{\texttt{4}} WARN\leavevmode\hphantom{..} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} & \cellcolor{bulmaRed} &
  \cellcolor{bulmaRed} & \cellcolor{bulmaRed}\\ \hline

  \protect\icircled{\texttt{3}} INFO\leavevmode\hphantom{..} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} &
  \cellcolor{bulmaRed} & \cellcolor{bulmaRed} \\ \hline

  \protect\icircled{\texttt{2}} DEBUG\leavevmode\hphantom{.} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaRed} \\ \hline

  \protect\icircled{\texttt{1}} TRACE\leavevmode\hphantom{.} & \cellcolor{bulmaGreen}
  & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen} &
  \cellcolor{bulmaGreen} & \cellcolor{bulmaGreen}\\ \hline

  \caption{
  \centering
  Correlation between logging level and logging message
  \newline
  \begin{tabular}{l}
    \fcolorbox{black}{bulmaGreen}{\rule{0pt}{6pt}\rule{6pt}{0pt}}\quad Logging message is recorded and saved \\
    \fcolorbox{black}{bulmaRed}{\rule{0pt}{6pt}\rule{6pt}{0pt}}\quad Logging message is completely ignored   \\
  \end{tabular}}
  \label{tbl:logging_levels}
\end{xltabular}