\chapter*{Abstract}
\label{abtract}
\addcontentsline{toc}{chapter}{Abstract}


Existing cluster architectures are primarily concerned with maximizing overall
performance and responsiveness, without consideration for the overall energy impact
and the many resources involved. Additionally, they are mostly composed of homogenous,
generally virtualized, machines that rely on modern and cutting-edge hardware components.
As a result, there is a need to reconsider the various methodologies and principles
involved, since it is becoming increasingly critical --- nearly unavoidable ---
to develop or transform cluster architectures orientated toward resource-awareness
and sustainability. The latter two requirements, where one does not negatively
influence the other, must be deemed as critical and significant as performance and
responsiveness. As a direct consequence, this new cluster architecture must retain
as much as possible of the previous architecture's usability and overall requirements
while being more mindful of the various components involved and its overall energy
consumption, achieving a more sustainable architecture not only suitable for the
present, but most importantly, for the future that is yet to come.

This document attempts to outline the many principles and methodologies that
should be considered when designing and developing a more resource-aware, sustainable,
and heterogeneous cluster architecture that is also simple to use and deploy. The
layout is intended to guide the reader through the different procedures,
decisions, and techniques that culminated in a final, realistic, and functional sustainable
cluster implementation. Every preceding chapter/section provides the theoretical
and practical foundation for the subsequent ones.

Before defining any kind of architecture or implementation, it is important to outline
the context in addition to the core principles around which the cluster is established.
A very dynamic and somewhat abstract context that is far from static and must
adapt/transform to the needs and requirements of the organization managing the
cluster. A collection of core principles that enhances and provides a basis for improved
sustainability, reusability, interoperability, and resource awareness while
minimizing waste as much as possible.

After the foundations, the theoretical architecture is depicted, which defines a
high-level overview of the whole cluster. This is a collection of all the involved
hardware and software components, along with their corresponding
responsibilities outlining how they manage and efficiently maintain the cluster,
both in terms of energy and resource awareness, and operationally as much as
possible and continuously. To communicate, the components must be connected, providing
a network that enables workload orchestration and cluster management. Three
network layers have been identified, each of which depicts a specific specialization:
K8s, Internal, and External; where the first layer is contained to the next layer,
and so on.

Finally, using the core principles and the theorized architecture, a
heterogeneous cluster has been implemented that attempts to be more resource-aware
and sustainable while being easy to use and fully functional like any other
existing cluster. The latter is identified by the name reCluster, which is
derived from the combination of the words recycling (since only reconditioned hardware
is being used) and cluster. The implementation defines the relationship between the
conceptual components and their software counterparts. Some of these
applications are partially pre-existing, but the two most important, the Server
and Cluster Autoscaler, necessitate, respectively, a fully custom implementation
built from the ground up and a custom Cloud Provider implementation that enables
the Cluster Autoscaler's core to communicate with the Server and vice versa.
Combining Server and Cluster Autoscaler allows the cluster architecture to
perform autoscaling procedures like upscaling and downscaling automatically. The
latter provides a completely autonomous, resource-aware, and sustainable cluster
by employing only the resources that are strictly required, preferring those
that are more sustainable than others. \\ %

A more sustainable future for computing environments is already here.